---
title: "Automated Reporting: DJIA"
author: "Joshua Boyer"
date: date()
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
#       User defined variables
# It is only necessary to alter the values contained here

# Dates: enter as yyyy-mm-dd
startDate = as.Date("2022-01-03")
endDate = as.Date("2022-01-31")

```



```{r, echo=FALSE, warning=FALSE}
# Import and clean raw data
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
rawData = read.csv("DJIA.csv")

#   Data cleaning
# Format dates to date type, format: yyyy-mm-dd
rawData$DATE = as.Date(rawData$DATE, format = "%Y-%m-%d")
periodLength = as.integer(endDate - startDate)

# Convert closing figures to numeric
rawData$DJIA = as.numeric(rawData$DJIA)

# Remove non-trading days included in the data
rawData = rawData[complete.cases(rawData), ]

# Validate that start and end dates are in data set
`%!in%` <- Negate(`%in%`)
if(startDate %!in% rawData$DATE){
    stop("The market was closed on the selected start date (holiday or weekend). Please select a new start date.")
    }

if(endDate %!in% rawData$DATE){
    stop("The market was closed on the selected end date (holiday or weekend). Please select a new end date.")
    }
```

```{r, echo=FALSE}
subsetData = subset.data.frame(rawData, 
                               rawData$DATE >= startDate & rawData$DATE <= endDate)
periodLength = as.integer(endDate - startDate)
totalTradingDays = length(subsetData$DJIA)


highestCloseValue = max(subsetData$DJIA)
lowestCloseValue = min(subsetData$DJIA)

highestCloseDate = subsetData$DATE[match(highestCloseValue, subsetData$DJIA)]
lowestCloseDate = subsetData$DATE[match(lowestCloseValue, subsetData$DJIA)]

# Calculate percent change from prior day
for (i in 2:totalTradingDays) {   
    subsetData$perChange[i] = round((subsetData$DJIA[i]/subsetData$DJIA[i-1]) - 1,
                                    digits = 5)
}

# Calculate absolute change from prior day
for (i in 2:totalTradingDays) {   
    subsetData$absChange[i] = round(subsetData$DJIA[i] - subsetData$DJIA[i-1],
                                    digits = 2)
}

perStats = c(round(mean(subsetData$perChange, na.rm = TRUE), digits = 5),
               round(sd(subsetData$perChange, na.rm = TRUE), digits = 5),
               round(max(subsetData$perChange, na.rm = TRUE), digits = 5),
               round(median(subsetData$perChange, na.rm = TRUE), digits = 5),
               round(min(subsetData$perChange, na.rm = TRUE), digits = 5)
               )

absStats = c(round(mean(subsetData$absChange, na.rm = TRUE), digits = 2),
               round(sd(subsetData$absChange, na.rm = TRUE), digits = 2),
               round(max(subsetData$absChange, na.rm = TRUE), digits = 2),
               round(median(subsetData$absChange, na.rm = TRUE), digits = 2),
               round(min(subsetData$absChange, na.rm = TRUE), digits = 2)
               )


statsTable = data.frame(perStats, absStats)
colnames(statsTable) = c("Percentage Change", "Absolute Change")
rownames(statsTable) = c("Mean", "Std. Dev.", "Max", "Median", "Min")
```

## Basic Information & Summary Statistics  
Reporting on Dow Jones closing figures from `r format(startDate, "%B %d %Y")` to `r format(endDate, "%B %d %Y")`.  
**Highest Close:** $`r formatC(highestCloseValue, format = "f", digits = 2, big.mark=",")` on `r format(highestCloseDate, "%B %d, %Y")`  
**Lowest Close:** $`r formatC(lowestCloseValue, format = "f", digits = 2, big.mark=",")` on `r format(lowestCloseDate, "%B %d, %Y")`  
  
`r knitr::kable(statsTable, align = "lll")`

```{r, echo = FALSE, fig.align = 'left'}
valueBarPlot = barplot(subsetData$DJIA,
                       ylim = c(0.85 * min(subsetData$DJIA), 
                                1.075 * max(subsetData$DJIA)),
                       xpd = FALSE,
                       xlab = "Date",
                       ylab = "Valuation ($)",
                       main = "Dow Jones Valuation",
                       col = "royalblue1")
box(bty="l")
axis(1, at = valueBarPlot, format(subsetData$DATE, format = "%m-%d"))
abline(a = mean(subsetData$DJIA), b = 0, lty = 2, lwd = 2)
legend(x = "topleft", legend = c("Period Average"), lty = 2, lwd = 2)
```
\newpage
```{r, echo = FALSE, fig.align = 'left'}
absLinPlot <- plot(subsetData$absChange,
                  xaxt = "n",
                  xlab = "",
                  ylab = "Change ($)",
                  main = "Absolute Change from Prior Day",
                  type = "b",
                  pch = 21,
                  bg = ifelse(subsetData$absChange < 0,'orangered','limegreen'),
                  col = "black"
                  )
abline(a = 0, b = 0, lty = 2)
```
Throughout this period, the market closed `r ifelse(mean(na.omit(subsetData$absChange)) > 0, "higher", "lower")` than the previous day more often than not. Overall, the market `r ifelse(subsetData$DJIA[length(subsetData$DJIA)] < subsetData$DJIA[1], "suffered losses", "realized gains")`, with the Dow ending the period $`r formatC(abs(subsetData$DJIA[length(subsetData$DJIA)] - subsetData$DJIA[1]), format = "f", digits = 2, big.mark=",")` `r  ifelse(subsetData$DJIA[length(subsetData$DJIA)] < subsetData$DJIA[1], "down", "up")`.



\newpage
```{r, echo = FALSE, fig.align = 'left'}
perBarPlot = barplot(subsetData$perChange,
                     xlab = "Date",
                     ylab = "Change (%)",
                     main = "Percentage Change from Prior Day",
                     col = ifelse(subsetData$perChange < 0,'orangered','limegreen')
                     )
axis(1, at = perBarPlot, format(subsetData$DATE, format = "%m-%d"))

daysTable = data.frame(sum(na.omit(subsetData$perChange > 0)),
                       sum(na.omit(subsetData$perChange < 0)))
colnames(daysTable) = c("Up Days", "Down Days")

```
`r knitr::kable(daysTable, align = "lll")`